---
title: "Comparisons"
author: "Insert authors here"
date: "`r format(Sys.time(), '%Y, %B %d')`"
output:
  html_document:
    toc: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
bibliography: ../manuscript/references.bib
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.align = "center")
knitr::opts_chunk$set(fig.height = 4)
knitr::opts_chunk$set(fig.width = 4)
t1 <- Sys.time()
```



Criteria

- Published
- Requires compilation
- Platforms / language
- Speed
- Works on FAST[AQ], gzipped


## Test data

The *Arabidopsis thaliana* genome (GCF_000001735.4_TAIR10.1_genomic.fna.gz) consists of 119,668,634 nucleotides arranged into 7 sequences with an N50 of 23,459,830 bp.
These 7 sequences include the mitochondrion (367,808 bp) and the chloroplast (154,478 bp).


```{r}
my_test_file <- "arab.fna.gz"
my_test_file2 <- sub("\\.gz$", "", my_test_file)

if( !file.exists(my_test_file) ){
  download.file("https://ftp.ncbi.nlm.nih.gov/genomes/all/GCF/000/001/735/GCF_000001735.4_TAIR10.1/GCF_000001735.4_TAIR10.1_genomic.fna.gz", destfile = my_test_file, method = "curl")
}
if( !file.exists(my_test_file2) ){
  my_cmd <- paste("gunzip -c ", my_test_file, " > ", my_test_file2, sep = "")
  system(my_cmd)
}

```



```{r}
summ_df <- data.frame( matrix(nrow = 8, ncol = 7) )
colnames(summ_df) <- c("Compiled", 
                       "FAST[AQ]",
                       "Gzip", 
                       "Tabular output",
                       "Sequence summary", "Graphics", "Time")
rownames(summ_df) <- c("nuccomp.py", "Quast", "SeqKit", 
                       "Seqstats", "sequence-stats",
                       "nuccomp.Rmd", "Quast_plots", "SeqKit_watch")

summ_df['nuccomp.py', 'Compiled'] <- "No"
summ_df['nuccomp.py', 'Gzip'] <- "Yes"
summ_df['nuccomp.py', 'Tabular output'] <- "Yes"
summ_df['nuccomp.py', 'Sequence summary'] <- "Yes"
summ_df['nuccomp.py', 'Graphics'] <- "No"

summ_df['nuccomp.Rmd', 'Compiled'] <- "No"
summ_df['nuccomp.Rmd', 'Gzip'] <- "Yes"
summ_df['nuccomp.Rmd', 'Tabular output'] <- "Yes"
summ_df['nuccomp.Rmd', 'Sequence summary'] <- "Yes"
summ_df['nuccomp.Rmd', 'Graphics'] <- "Yes"

summ_df['Quast', 'Compiled'] <- "No"
summ_df['Quast', 'Gzip'] <- "Yes"
summ_df['Quast', 'Tabular output'] <- "Yes"
summ_df['Quast', 'Sequence summary'] <- "No"
summ_df['Quast', 'Graphics'] <- "No"

summ_df['Quast_plots', 'Compiled'] <- "No"
summ_df['Quast_plots', 'Gzip'] <- "Yes"
summ_df['Quast_plots', 'Tabular output'] <- "Yes"
summ_df['Quast_plots', 'Sequence summary'] <- "No"
summ_df['Quast_plots', 'Graphics'] <- "Yes"

summ_df['SeqKit', 'Compiled'] <- "No"
summ_df['SeqKit', 'Gzip'] <- "Yes"
summ_df['SeqKit', 'Tabular output'] <- "Yes"
summ_df['SeqKit', 'Sequence summary'] <- "No"
summ_df['SeqKit', 'Graphics'] <- "No"

summ_df['SeqKit_watch', 'Compiled'] <- "No"
summ_df['SeqKit_watch', 'Gzip'] <- "Yes"
summ_df['SeqKit_watch', 'Tabular output'] <- "No"
summ_df['SeqKit_watch', 'Sequence summary'] <- "No"
summ_df['SeqKit_watch', 'Graphics'] <- "Yes"

```



## nuccomp.py

Versioning: https://github.com/knausb/nuccomp

Website: https://github.com/knausb/nuccomp

Platform: Unix, Windows

```{r}
my_cmd <- paste("~/gits/nuccomp/python/nuccomp.py ", my_test_file, "", sep = "")
my_time <- system.time( system(my_cmd) )
#my_time

my_nuccomp <- read.csv("arab_nuccomp.csv")
knitr::kable(my_nuccomp[, 1:8], caption = "Example output produced by nuccomp.py.")

summ_df['nuccomp.py', 'Time'] <- my_time['elapsed']
#summ_df

```



```{r}
file.remove("./nuccomp.Rmd")
file.copy(from = "../R/nuccomp.Rmd", to = "./nuccomp.Rmd")

my_cmd <- paste("perl -pi -e 's/my_infile <- \"S288C_reference_sequence_R64-2-1_20150113.fsa.gz\"/my_infile <- \"", my_test_file, "\"/g' nuccomp.Rmd", sep = "")
system( my_cmd )

my_cmd <- "Rscript -e \"rmarkdown::render('nuccomp.Rmd')\""
my_time <- system.time( system( my_cmd ) )

#rmarkdown::render('nuccomp.Rmd')
#my_time <- system.time( rmarkdown::render('nuccomp.Rmd') )
summ_df['nuccomp.Rmd', 'Time'] <- my_time['elapsed']

```




## Quast

Citation: @gurevich2013quast

Versioning: https://github.com/ablab/quast/

Website: https://quast.sourceforge.net/docs/manual.html

Platform: Unix



```{r}
my_cmd <- paste("~/gits/quast/quast.py -v", sep = "")
my_time <- system.time( system(my_cmd) )

my_cmd <- paste("~/gits/quast/quast.py --threads 1 --large --silent --no-plots --no-html ", my_test_file, sep = "")
my_time <- system.time( system(my_cmd) )

summ_df['Quast', 'Time'] <- my_time['elapsed']
#summ_df


#my_quast <- read.table("quast_results/results_2024_06_13_02_07_08/report.tsv", sep = "\t", header = T)
my_quast <- read.table("quast_results/latest/report.tsv", sep = "\t", header = T)
knitr::kable(my_quast, format.args = list(big.mark = ","), caption = "Example output from file produced by Quast.")

```


```{r}
my_cmd <- paste("~/gits/quast/quast.py -v", sep = "")
my_time <- system.time( system(my_cmd) )

my_cmd <- paste("~/gits/quast/quast.py --threads 1 --large --silent ", my_test_file, sep = "")
my_time <- system.time( system(my_cmd) )

summ_df['Quast_plots', 'Time'] <- my_time['elapsed']
#summ_df

#my_quast <- read.table("quast_results/results_2024_06_13_02_07_08/report.tsv", sep = "\t", header = T)
my_quast <- read.table("quast_results/latest/report.tsv", sep = "\t", header = T)
knitr::kable(my_quast, format.args = list(big.mark = ","), caption = "Example output from file produced by Quast.")
```




## SeqKit

Citation: @shen2024seqkit2

Versioning: https://github.com/shenwei356/seqkit

Website: https://bioinf.shenwei.me/seqkit/

Platform: Unix, Windows


Output is to standard out, but includes a '--tabular' flag I used to read directly into R, but could also be redirected to a file on platforms that support this.

Includes modules that produce graphics.


```{r}
my_cmd <- "~/bin/seqkit/seqkit version"
my_time <- system.time( system(my_cmd) )


#my_cmd <- paste("~/bin/seqkit/seqkit seq ", my_test_file, " --out-file seqkit_out.txt stats", sep = "")
my_cmd <- paste("~/bin/seqkit/seqkit stats ", my_test_file, " --tabular ", sep = "")
my_time <- system.time( SKout <- system(my_cmd, intern = TRUE) )
#my_time

#print(SKout)
SKout <- read.table(text=SKout, sep = "\t", header = T)
knitr::kable(SKout, format.args = list(big.mark = ","))


summ_df['SeqKit', 'Time'] <- my_time['elapsed']
#summ_df

# my_cmd <- paste("~/bin/seqkit/seqkit sum ", my_test_file, "", sep = "")
# my_time <- system.time( SKout <- system(my_cmd, intern = TRUE) )
# my_time
```


```{r}
my_cmd <- paste("~/bin/seqkit/seqkit watch --bins 7 --fields ReadLen ", my_test_file, " -O len.png", sep = "")
my_time <- system.time( SKout <- system(my_cmd, intern = TRUE) )
#my_time
summ_df['SeqKit_watch', 'Time'] <- my_time['elapsed']

```






## Seqstats


Citation: none

Versioning: https://github.com/clwgg/seqstats

Website: https://github.com/clwgg/seqstats

Platform: Unix


C/make

Requires compilation and make.

```
$ make
make target 'klib/kseq.h', needed
```

```
$ make
make: *** No rule to make target 'klib/kseq.h', needed by 'seqstats'.  Stop.
```

Writes to standard out.
I was able to capture this but do not consider it 'machine friendly'.
The 'N 50' record included '\t\t' where other records included '\t'.
This may be easy for some users to sanitize, but may present a barier to use to other users.


```{r}
#my_cmd <- "~/gits/seqstats/seqstats"

my_cmd <- paste("~/gits/seqstats/seqstats ", my_test_file, "", sep = "")
my_time <- system.time( seqs_out <- system(my_cmd, intern = TRUE) )
#my_time

summ_df['Seqstats', 'Compiled'] <- "Yes"
summ_df['Seqstats', 'Gzip'] <- "Yes"
summ_df['Seqstats', 'Tabular output'] <- "No"
summ_df['Seqstats', 'Sequence summary'] <- "No"
summ_df['Seqstats', 'Graphics'] <- "No"
summ_df['Seqstats', 'Time'] <- my_time['elapsed']

seqs_out <- sub("\t\t", "\t", seqs_out)
seqs_out <- read.table( text = seqs_out, sep = "\t", header = F)
#read.table( text = seqs_out, sep = "\t", header = F)

knitr::kable(seqs_out, format.args = list(big.mark = ","), caption = "Example output produced by seqstats.")

```





## seqstats

https://github.com/drostlab/seqstats

No citation.



## sequence-stats

Citation: none

Versioning: https://github.com/raymondkiu/sequence-stats

Website: https://github.com/raymondkiu/sequence-stats

Platform: Unix

Does not handle gzipped data.

Output sent to standard out was successfully captured in R and treated as colon delimited tabular data.

Uses awk and BASH.


```{r}

my_cmd <- "~/gits/sequence-stats/src/sequence-stats -v"
my_time <- system.time( system(my_cmd) )

my_cmd <- paste("~/gits/sequence-stats/src/sequence-stats -a ", my_test_file2, "", sep = "")
my_time <- system.time( my_sstat <- system(my_cmd, intern = TRUE) )
#my_time
my_sstat <- read.table(text = my_sstat, sep  =":", header = TRUE)
knitr::kable(my_sstat, format.args = list(big.mark=",", scientific=F))

my_cmd <- paste("~/gits/sequence-stats/src/sequence-stats -c ", my_test_file2, "", sep = "")
my_time2 <- system.time( my_sstat <- system(my_cmd, intern = TRUE) )
my_sstat <- read.table(text = my_sstat, sep  ="\t", header = FALSE)
knitr::kable(my_sstat, format.args = list(big.mark=",", scientific=F))
#
my_time <- my_time + my_time2

my_cmd <- paste("~/gits/sequence-stats/src/sequence-stats -b ", my_test_file2, "", sep = "")
#my_time <- system.time( system(my_cmd) )
#my_time
#print(my_time[1:3])
#my_time['elapsed']

summ_df['sequence-stats', 'Compiled'] <- "No"
summ_df['sequence-stats', 'Gzip'] <- "No"
summ_df['sequence-stats', 'Tabular output'] <- "Yes"
summ_df['sequence-stats', 'Sequence summary'] <- "Yes"
summ_df['sequence-stats', 'Graphics'] <- "No"
summ_df['sequence-stats', 'Time'] <- my_time['elapsed']
#summ_df
```


## Summary


```{r}
names(summ_df)[ names(summ_df) == "Time" ] <- "Seconds"
summ_df$Seconds <- round(summ_df$Seconds, digits = 3)
summ_df <- data.frame(
  Software = row.names(summ_df),
  summ_df
)


knitr::kable(summ_df, row.names = FALSE)



write.table(summ_df, file = "comparison_summary.csv",
            sep = ",", row.names = FALSE, col.names = TRUE)

```



## Clean up


```{r}
# if( file.exists(my_test_file) ){
#   file.remove(my_test_file)
# }
```


```{r}
t99 <- Sys.time()
t99 - t1
```

```{r}
sessionInfo()
```


## References
